\textsf{Are large language models superhuman chemists?}

### 2024-10-25 src/tex/appendix.tex
feat: incorporate feedback from Santiago

\Cref{fig:performance_corpus_and_tiny} shows the performance of the models on the \chembench corpus and the \chembenchmini subset. The relative performance difference between both is actually quite similar across most models. This makes \chembenchmini subset a reliable subset for human baseline comparison and particularly valuable for rapid prototyping and initial model assessment phases.

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/corpus_human_comparison.pdf}
    \caption{\textbf{
        Performance of the models on the \chembench corpus and \chembenchmini subset.} The bar plot shows the fraction of questions that were answered completely correctly, highlighting the relative performance of different models on both \chembench corpus and the \chembenchmini subset.
        We see that the model ranking remains fairly consistent across both sets}  
    }
    \label{fig:performance_corpus_and_tiny}
    \script{corpus_humanset_performance.py}
\end{figure}

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of NMR signals) is not correlated with the complexity of the molecules. For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
    \script{correlate_with_molecule_features.py}
    \label{fig:correlation_plot_is_number_nmr_peaks_complexity}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_score_distribution.pdf}
    \script{plot_human_score_distribution.py}
    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one. A score of one would mean that a volunteer answered all questions correctly. A score of zero would mean that no question was answered correctly.}
    \label{fig:human_score_distribution}
\end{figure}

>>>>

@@

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/correlation_plot_is_number_nmr_peaks_complexity.pdf}
    \caption{\textbf{Dependence of the mean absolute error in predicting the number of NMR signals on the Böttcher complexity of the molecules.} The complexity measure proposed by \textcite{B_ttcher_2016} is an information-theoretic additive measure of compound complexity that follows chemical intuitions.
    The plot shows that for the \glspl{llm}, the predictive performance (measured as the mean absolute error in the prediction of the number of \gls{nmr} signals) is not correlated with the complexity of the molecules (that is, molecules tend to not to be able to predict the number of \gls{nmr} signals irregardless of molecular complexity). For inference based on reasoning, one would expect that the complexity of the molecule is a good predictor of the difficulty of the question.}
    \script{correlate_with_molecule_features.py}
    \label{fig:correlation_plot_is_number_nmr_peaks_complexity}
\end{figure}

@@

\begin{figure}[htb]
    \centering
    \includegraphics{figures/human_score_distribution.pdf}
    \script{plot_human_score_distribution.py}
    \caption{\textbf{Distribution of human scores.} The histogram and kernel density estimates show the fraction of questions answered correctly by the human volunteers.
    Since the best possible score for each question is one and the worst possible score is zero, the values on this plot are between zero and one. A score of one would mean that a volunteer answered all questions correctly. A score of zero would mean that no question was answered correctly.
    We find that the scores for the questions that the human volunteers answered with tools are generally lower than the scores for the questions that the human volunteers answered without tools.
    }
    \label{fig:human_score_distribution}
\end{figure}

### 2024-10-25 src/tex/ms.tex
feat: incorporate feedback from Santiago

Chemists and materials scientists have quickly caught on to the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language.}\autocite{White_2023}
This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions,\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine, alampara2024mattextlanguagemodelsneed} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lala2023paperqa, caufield2023structured, gupta2022discomat, schillingwilhelmi2024textinsightlargelanguage, skarlinski2024language} or to even prototype systems that can autonomously perform reactions in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}

@@

 \begin{figure}[!h]
     \centering
     \includegraphics[width=\textwidth]{figures/model_confidence_performance.pdf}
     \caption{\textbf{Calibration and distribution of confidence estimates.} For this analysis, we used verbalized confidence estimates from the model. We prompted the models to return a confidence score on an ordinal scale to obtain those estimates.
     The line plot shows the average fraction of correctly answered questions for each confidence level. The bar plot shows the distribution of confidence estimates. A confidence estimate would be well-calibrated if the average fraction of correctly answered questions increases with the confidence level.
     We find that most models are not well-calibrated and provide misleading confidence estimates.
     }
     \label{fig:confidence_vs_performance}
     \script{confidence_estimate.py}
 \end{figure}

@@

We employ distinct prompt templates tailored for completion and instruction-tuned models to maintain consistency with the training.
As explained below, we impose constraints on the models within these templates to receive responses in a specific format so that robust, fair, and consistent parsing can be performed.
Certain models are trained with special annotations and \LaTeX\xspace syntax for scientific notations, chemical reactions, or symbols embedded within the text.
For example, all the SMILES representations are encapsulated within \texttt{[START\_SMILES][\textbackslash END\_SMILES]} in Galactica\autocite{taylor2022galactica}.
Our prompting strategy consistently adheres to these details in a model-specific manner by post-processing \LaTeX\xspace syntax, chemical symbols, chemical equations, and physical units (by either adding or removing wrappers).
This step can be easily customized in our codebase.

@@

\section*{Acknowledgements}
This work was supported by the Carl Zeiss Foundation, a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.
We also thank Stability.AI for the access to its HPC cluster. In addition, M.S-W.'s work was supported by Intel and Merck via the AWASES programme.
K.M.J.\ is part of the NFDI consortium FAIRmat funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – project 460197019.
K.M.J.\ thanks FutureHouse (a non-profit research organization supported by the generosity of Eric and Wendy Schmidt) for supporting \PaperQATwo run by access to the \gls{api}.
Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.
A.A.\ gratefully acknowledges financial support for this research by the Fulbright U.S. Student Program, which is sponsored by the U.S.\ Department of State and German-American Fulbright Commission. Its contents are solely the responsibility of the author and do not necessarily represent the official views of the Fulbright Program, the Government of the United States, or the German-American Fulbright Commission.

>>>>

Chemists and materials scientists have quickly caught on to the mounting attention given to \glspl{llm}, with some voices even suggesting that \enquote{the future of chemistry is language.}\autocite{White_2023}
This statement is motivated by a growing number of reports that use \glspl{llm} to predict properties of molecules or materials,\autocite{jablonka202314, jablonka2024leveraging, xie2024fine, liao2024words, zhang2024chemllm, zhong2024benchmarking}  optimize reactions,\autocite{ramos2023bayesian, kristiadi2024sober}  generate materials,\autocite{rubungo2023llm, flam2023language, gruver2024fine, alampara2024mattextlanguagemodelsneed} extract information,\autocite{Patiny_2023, Dagdelen_2024, Zheng_2024, lala2023paperqa, caufield2023structured, gupta2022discomat, schillingwilhelmi2024textinsightlargelanguage, skarlinski2024language} or to even prototype systems that can autonomously perform experiments in the physical world based on commands provided in natural language.\autocite{bran2023chemcrow, Boiko_2023, darvish2024organa}

@@

 \begin{figure}[!h]
     \centering
     \includegraphics[width=\textwidth]{figures/model_confidence_performance.pdf}
     \caption{\textbf{Calibration and distribution of confidence estimates.} For this analysis, we used verbalized confidence estimates from the model. We prompted the models to return a confidence score on an ordinal scale to obtain those estimates.
     The line plot shows the average fraction of correctly answered questions for each confidence level. The bar plot shows the distribution of confidence estimates.
     A confidence estimate would be well-calibrated if the average fraction of correctly answered questions increases with the confidence level.      The dashed black line indicates this ideal behavior, which would be monotonically increasing correctness with higher
     levels of confidence.
     We find that most models are not well-calibrated and provide misleading confidence estimates.
     }
     \label{fig:confidence_vs_performance}
     \script{confidence_estimate.py}
 \end{figure}

@@

We employ distinct prompt templates tailored for completion and instruction-tuned models to maintain consistency with the training.
As explained below, we impose constraints on the models within these templates to receive responses in a specific format so that robust, fair, and consistent parsing can be performed.
Certain models are trained with special annotations and \LaTeX\xspace syntax for scientific notations, chemical reactions, or symbols embedded within the text.
For example, all the \gls{smiles} representations are encapsulated within \texttt{[START\_SMILES][\textbackslash END\_SMILES]} in Galactica\autocite{taylor2022galactica}.
Our prompting strategy consistently adheres to these details in a model-specific manner by post-processing \LaTeX\xspace syntax, chemical symbols, chemical equations, and physical units (by either adding or removing wrappers).
This step can be easily customized in our codebase and we provide presets for the models we evaluated.

@@

\section*{Acknowledgements}
This work was supported by the Carl Zeiss Foundation, and a \enquote{Talent Fund} of the \enquote{Life} profile line of the Friedrich Schiller University Jena.

In addition, M.S-W.'s work was supported by Intel and Merck via the AWASES programme.

Parts of A.M.'s work was supported as part of the \enquote{SOL-AI} project funded by the Helmholtz Foundation model initative.

K.M.J.\ is part of the NFDI consortium FAIRmat funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – project 460197019.

K.M.J.\ thanks FutureHouse (a non-profit research organization supported by the generosity of Eric and Wendy Schmidt) for supporting \PaperQATwo runs via access to the \gls{api}. We also thank Stability.AI for the access to its HPC cluster.

A.A.\ gratefully acknowledges financial support for this research by the Fulbright U.S. Student Program, which is sponsored by the U.S.\ Department of State and German-American Fulbright Commission. Its contents are solely the responsibility of the author and do not necessarily represent the official views of the Fulbright Program, the Government of the United States, or the German-American Fulbright Commission.
